Type cones of $\b{g}$-vector fans

### 2019-02-04 thirdPaper/draft/typeConeAssociahedra.tex
More changes to section 3

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Relations for~$\b{g}$-vectors in~$2$-Calabi--Yau triangulated categories}
%%%%%
\subsection{Setting}
Let~$\field$ be a field.  Let~$\cat$ be a~$\field$-linear triangulated category with suspension functor~$\susp$.  We fix a collection~$\ind(\cat)$ of representatives of isomorphism classes of indecomposable objects of~$\cat$. We will assume the following:
\begin{itemize}
 \item $\cat$ is essentially small (in particular,~$\ind(\cat)$ is a set);
 \item $\cat$ is~$\Hom{}$-finite: for each pair of objects~$X$ and~$Y$, the~$\field$-vector space~$\cat(X,Y)$ is finite-dimensional;
 \item $\cat$ is Krull--Schmidt: the endomorphism algebra of any indecomposable object is local;
 \item $\cat$ is~$2$-Calabi--Yau: for each pair of objects~$X$ and~$Y$, there is an isomorphism of bifunctors
 \[
  \cat(X,\susp Y) \to D\cat(Y,\susp X)
 \]
 where~$D=\Hom{\field}(-,\field)$ is the usual duality of vector spaces;
 \item $\cat$ contains a basic cluster-tilting object~$T=\bigoplus_{i=1}^n T_i$:
 \[
  \textrm{for any object~$X$,} \quad \cat(T, \susp X) = 0 \quad \textrm{if and only if} \quad X\in \add(T),
 \]
 where~$\add(T)$ is the smallest additive full subcategory of~$\cat$ containing the~$T_i$'s and closed under isomorphisms.
\end{itemize}

@@

\begin{theorem}\label{theo::relations-g-vecteurs}
 Let~$\cat$ be a category satisying the hypotheses of \ref{} \pg{add ref}.  Then~$\cat$ has only finitely many isomorphism classes of indecomposable objects if and only if the set~\[L:=\big\{\ell_X \ | \ X\in \ind(\cat)\setminus \add(\susp T) \big\}\] generates the kernel of~$\b g:\Ksp(\cat) \to \Kg(\cat)$. 
 In this case, the set~$L$ is a basis of the kernel of~$\b g$, and \pg{add formula}, for any $x\in \ker(\b g)$, we have that
 \[
  x= \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
 \]
\end{theorem}

@@

\begin{lemma}
 If~$X$ or~$Y$ lie in~$\add(\susp T)$, then~$\langle X, Y \rangle = 0$.
\end{lemma}
\begin{proof}
 This is because~$F\susp T = 0$.
\end{proof}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Relations for~$\b{g}$-vectors in~$2$-Calabi--Yau triangulated categories}
%%%%%
\subsection{Setting}\label{sect::setting}
Let~$\field$ be a field.  Let~$\cat$ be a~$\field$-linear triangulated category with suspension functor~$\susp$.  We fix a collection~$\ind(\cat)$ of representatives of isomorphism classes of indecomposable objects of~$\cat$. We will assume the following:
\begin{itemize}
 \item $\cat$ is essentially small (in particular,~$\ind(\cat)$ is a set);
 \item $\cat$ is~$\Hom{}$-finite: for each pair of objects~$X$ and~$Y$, the~$\field$-vector space~$\cat(X,Y)$ is finite-dimensional;
 \item $\cat$ is Krull--Schmidt: the endomorphism algebra of any indecomposable object is local;
 \item $\cat$ is~$2$-Calabi--Yau: for each pair of objects~$X$ and~$Y$, there is an isomorphism of bifunctors
 \[
  \cat(X,\susp Y) \to D\cat(Y,\susp X)
 \]
 where~$D=\Hom{\field}(-,\field)$ is the usual duality of vector spaces;
 \item $\cat$ contains a basic cluster-tilting object~$T=\bigoplus_{i=1}^n T_i$:
 \[
  \textrm{for any object~$X$,} \quad \cat(T, \susp X) = 0 \quad \textrm{if and only if} \quad X\in \add(T),
 \]
 where~$\add(T)$ is the smallest additive full subcategory of~$\cat$ containing the~$T_i$'s and closed under isomorphisms.
\end{itemize}

@@

\begin{theorem}\label{theo::relations-g-vecteurs}
 Let~$\cat$ be a category satisying the hypotheses of Section \ref{sect::setting}.  Then~$\cat$ has only finitely many isomorphism classes of indecomposable objects if and only if the set~\[L:=\big\{\ell_X \ | \ X\in \ind(\cat)\setminus \add(\susp T) \big\}\] generates the kernel of~$\b g:\Ksp(\cat) \to \Kg(\cat)$. 
 In this case, the set~$L$ is a basis of the kernel of~$\b g$, and for any $x\in \ker(\b g)$, we have that
 \[
  x= \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
 \]
\end{theorem}

@@

\begin{lemma}\label{lemm::bilinear-form-suspention-T}
 If~$X$ or~$Y$ lie in~$\add(\susp T)$, then~$\langle X, Y \rangle = 0$.
\end{lemma}
\begin{proof}
 This is because~$F\susp T = 0$.
\end{proof}

@@

\begin{lemma}\label{lemm::coefficients}
 Let~$x\in \Ksp(\cat)$, and write~\[x=\sum_{A\in \ind(\cat)} \lambda_A [A].\]  Then for any~$A\in\ind(\cat)\setminus \add(\susp T)$, we have that
 \[
  \lambda_A = \langle \ell_A, x \rangle = \langle x, r_A \rangle.
 \]
\end{lemma}
\begin{proof}
 Let~$B\in \ind(\cat) \setminus \add(\susp T)$.  Applying Lemma~\ref{lemm::bilinear-form}, we get that
 \[
  \langle \ell_B, x \rangle = \langle \ell_B, \sum_{A\in \ind(\cat)} \lambda_A [A] \rangle = \sum_{A\in \ind{\cat}} \lambda_A \langle \ell_B, [A] \rangle = \lambda_A.
 \]
 The equality~$\lambda_A = \langle x, r_A\rangle$ is proved in a similar way.
\end{proof}

\begin{corollary}\label{lemm::test-susp-T}
 Let~$x\in \Ksp(\cat)$.  Then the following are equivalent.
 \begin{enumerate}
  \item $x\in \Ksp(\add(\susp T))$;
  \item $\langle x, [A]\rangle = 0$ for all~$A\in \ind(\cat)$;
  \item $\langle [A], x \rangle = 0$ for all~$A\in \ind(\cat)$.
 \end{enumerate}
\end{corollary}
\begin{proof}
 We will only proof that (1) is equivalent to (2); the proof that (1) is equivalent to (3) is similar.

 Assume that (2) holds.  Then, by Lemma~\ref{lemm::coefficients}, we have that
 \[
  x = \sum_{A\in \ind(\cat)\setminus \add(\susp T)} \langle \ell_A, x \rangle + \sum_{i=1}^n \lambda_{\susp T_i} [\susp T_i] = \sum_{i=1}^n \lambda_{\susp T_i} [\susp T_i].
 \]
 Thus~$x\in \Ksp(\add(\susp T))$, and (1) holds.

 Assume now that (1) holds.  Then~$\langle x, [A]\rangle = 0$ for any~$A$ by Lemma~\ref{lemm::bilinear-form-suspention-T}.  Thus (2) holds.
\end{proof}

\begin{proposition}\label{prop::free-set}
 \begin{enumerate}
  \item The set~$\{[\ell_A] \ | \ A\in \ind(\cat) \setminus \add(\susp T) \} \cup \{[\susp T_i] \ | \ i=1, \ldots, n\}$ is free in~$\Ksp(\cat)$.
  \item The set~$\{[r_A] \ | \ A\in \ind(\cat) \setminus \add(\susp T) \} \cup \{[\susp T_i] \ | \ i=1, \ldots, n\}$ is free in~$\Ksp(\cat)$.
 \end{enumerate}
\end{proposition}
\begin{proof}
 We only prove (1); the proof of (2) is similar.  Assume that~
 \[
  x= \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \lambda_A \ell_A + \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
 \]
 Then~$\langle x, [A] \rangle = 0$ for all~$A\in\ind(\cat) \setminus \add(\susp T)$.  But~$\langle x, [A] \rangle = \lambda_A$ by Lemma~\ref{lemm::bilinear-form}.  Thus~
 \[
  x = \sum_{i=1}^n \lambda_i [\susp T_i] = 0.
 \]
 But the~$[\susp T_i]$ are linearly independent in~$\Ksp(\cat)$.  Thus~$\lambda_i = 0$ for all~$i\in \{1, \ldots, n\}$.  This finishes the proof.
\end{proof}

\begin{proposition}
 Assume that~$\ind(\cat)$ is finite.  Then the set~$\{[\ell_A] \ | \ A\in \ind(\cat) \setminus \add(\susp T) \}$ is a basis of the kernel of~$\b g$.  Moreover, for any~$x\in \ker \b g$, we have that
 \[
  x= \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
 \]
\end{proposition}
\begin{proof}
By Proposition~\ref{prop::free-set}, the set is free.  Let~$x\in \ker \b g$.  Consider the element
\[
 z = x- \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
Then for any~$B\in \ind{\cat}$, we have that
\[
 \langle  z, [B] \rangle = \Big\langle x- \sum_{A\in \ind(\cat) \setminus \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A, [B] \Big\rangle = \langle x, [B] \rangle - \langle x, [B] \rangle = 0, 
\]
where the second equality is obtained by using Lemma~\ref{lemm::bilinear-form}.  By Lemma~\ref{lemm::test-susp-T}, this implies that~$z\in \Ksp(\add(\susp T))$.  Since~$z\in \ker(\b g)$ and since~$\b g$ is injective on~$\Ksp(\add (\susp T))$, we get that~$z=0$.  This finishes the proof.
\pg{We need a lemma saying that g is injective on K(add(susp T))}
\end{proof}

All that remains is to prove the converse in the statement of Theorem~\ref{theo::relations-g-vecteurs}.

\begin{proposition}
 Assume that the set~$\{[\ell_A] \ | \ A\in \ind(\cat) \setminus \add(\susp T) \}$ is a basis of the kernel of~$\b g$. Then~$\ind(\cat)$ is finite.
\end{proposition}
\begin{proof}

\end{proof}
